\chapter{Introduction}


Wireless Sensor Networks consist of a large number of sensor nodes, which are battery-powered, low energy consuming devices. Wireless Sensor Networks (WSNs) are used for applications such as monitoring (e.g., pollution prevention~\cite{application2}, structures and buildings health), event detection (e.g. fire hazard ~\cite{application5}) and target tracking (e.g., surveillance~\cite{application3}). These devices perform three basic tasks:
\begin{itemize}
\item Sample a physical quantity from the surrounding environment
\item Process and buffer the acquired data
\item Transfer the data through wireless communications to a data collection point called sink node or base station.
\end{itemize}
The sensors, as soon as deployed on the field, identify their adjacent sensors (neighbours), and are able to form an ad-hoc network. If this network is connected (each sensor is able to send data to another sensor on the field through one or multiple hops), then it is called a dense sensor network. Such a network is used as basic infrastructure for routing sensor-acquired data to the sink node (Base Station) using WSN routing protocols~\cite{routingSurvey}. Occasionally, sink node may send data to all sensor nodes as well. 

However, dense networks are seldom realizable in practice. Usually, sensors are spread (air dropped) over a large geographical area resulting in a sparse network. This kind of networks are useful for applications (e.g., environmental monitoring applications~\cite{sparseEx}) where fine-grained sensing is not required. In sparse sensor networks, islands of sensor sub-networks are formed (WSN islands~\cite{intro2}), which cannot communicate to each other without external help. Data collection from this kind of network, without deploying additional sensors can be done using:
\begin{itemize}
\item Long range relay nodes~\cite{relayNodes}
  
Long range and high power nodes, whose main purpose is to establish connectivity among such WSN islands, and their connectivity to the base station by relaying/receiving data to/from sensors out of range of the local island. Their placement in a sensor network is a well studied problem.
\item Mobile data collectors (MDCs)~\cite{intro3} or Mobile Ubiquitous LAN Extensions (MULEs)~\cite{intro1}: 
  
Controlled mobile robots acting as mobile sinks/Base stations/relay nodes. These can be programmed to tour any set of locations so as to collect data from the sensors at those positions,] and relay the same to base station directly or through another MULE. Since the sensors have to wait for the MULE to arrive in range
\begin{enumerate}
\item Sensors may be required to have a battery powered passive wake-up radio device~\cite{intro4}, which can be triggered by a nearby MULE. MULE will then be able to collect data from the sensor, and forward the collected data 
when it is in proximity of the base station (or another MULE assigned to relay to the base station~\cite{sim2}). %it will forward this collected data.

\item Sensors wake-up and sleep operation can be controlled at the MAC layer~\cite{dutyCycle1} \cite{dutyCycle2}. In this approach, sensors periodically sleep and wakeup to save power. Data transmission occurs only when the MULE is in range, and the sensor is awake.
\end{enumerate}
\end{itemize}

The MULE approach naturally saves energy of the sensors leading to prolonging network lifetime. The use of MULEs limits sensor communication to one hop only, and relieves sensors from perform aggregation or forwarding of data of other sensors. However, the main disadvantage in this approach is increase in latency of data transfer from sensors to Base Station. This may have negative impact onperformance of some applications~\cite{application4}. This is the reason that the studies on the problems of MULE path optimization and job scheduling are extensive in literature~\cite{muleSurvey}.

\section{Main Contribution of the Work}
In this thesis our investigation is centered around study of techniques and heuristics used for the use of MULEs in data collection from sensors. The main contribution of our work is to compare existing heuristics and propose new heuristics for data collection through MULEs. The usual approach in this direction is to make use of multiple MULEs by partitioning the network into roughly "equal" parts (in terms of TSP tour time of that partition, together with the time taken to collect data from sensors in that partition), and each partition is assigned to a MULE. Our work consisters the latency constraint of the sensor network application as input and computes the number of MULEs required along with their trajectories. We assume MULEs can move freely in the plane of the sensor network, and they have identical, constant speed.

\section{Summary of the Work}
This thesis has been organized into five chapters including the 
introductory chapter. In Chapter~\ref{chapter:2}, we first deal with the components of the general problem of data MULE scheduling, and with the aspects that we are concerned, namely, {\em path selection}. Next, we introduce location nodes, location graphs and propose heuristics to compute them. We also provided justification of our choice of heuristic with performance measures. In Chapter~\ref{chap:steiner}, we describe our main heuristic for path selection and also provide pseudocode for the same. 
In Chapter~\ref{chapter:5}, we present the results of simulation studies conducted to examine the performance the proposed heuristics for data collection using multiple MULEs. Finally, 
Chapter~\ref{chap:concl} contains two anticipated directions in which this work may be continued. 
The first is the addition of capability of MULEs to avoid obstacles, and the second is incorporating optimization of TSP tours for MULEs in the proposed heuristics. A MULE need not travel to exact location of a sensor node in order to collect data. Data transfer can be staged when MULE enters into the communication range of a sensor. So, in the proposed heuristic, it suffices to obtain a TSP tour for a MULEs that passes through communication neighbourhood of sensors instead of exact locations. %In the problem of TSP tours in presence of neighbourhoods, in order to visit a node, the salesman (MULE) need not travel to the exact location of the node. Itsuffices if the MULE just enters a given neighbourhood of the node. %For example, a MULE may just enter a sensor node's communication area instead of travelling to the sensor position for data collection.
